A Decision Procedure for (Co)datatypes in SMT Solvers

### 2014-12-09 tex/codatadp.tex
Example run of model generation.

\rn{Base}
\(
\inferrule{
  %x : \tau
  %\quad
  %\tau \in \Codata
  %\quad
  \ec{x} \mapsto v_{\ec{x}} \in \Val
  %\quad
  %\ec{x} \text{ ctr-free}
  \quad
  c \not\sim c' \text{ for all } c' \in \vrange(\Val)
}{
  \Val := \Val \{ v_{\ec{x}} \mapsto c \}
}
\)

\rem{define $\sim$}
We assume these rules are applied to completion in the order in which they are written,
that is, \rn{Ind} is applied exhaustively before applying \rn{Base}.
%the rule \rn{D-Ind} is applied to completion before applying \rn{D-Base}.
In the rule \rn{Ind}, we write $FV( \Val )$ to denote the free variables occurring in $\Val$.
Notice that this excludes variables beneath $\mu$-binders.
It is clear that after applying these rules to completion, $\Val$ is a mapping from equivalence classes to ground values.

>>>>

\rn{Base}
\(
\inferrule{
  %x : \tau
  %\quad
  %\tau \in \Codata
  %\quad
  v_{\ec{x}} \in FV( \Val )
  %\quad
  %\ec{x} \text{ ctr-free}
  \quad
  c \not\sim c' \text{ for all } c' \in \vrange(\Val)
}{
  \Val := \Val \{ v_{\ec{x}} \mapsto c \}
}
\)

\rem{define $\sim$}
We assume that first \rn{Ind} is applied exhaustively, after which \rn{Base} is applied exhaustively.
%the rule \rn{D-Ind} is applied to completion before applying \rn{D-Base}.
In the rule \rn{Ind}, we write $FV( \Val )$ to denote the free variables occurring in $\Val$.
Notice that this excludes variables beneath $\mu$-binders.

As an example,
say that $\Fc$ contains equivalence classes $\ec{x}$ and $\ec{y}$ where $\const{C}(z,x) \in \ec{y}$ and $\const{C}(w,y) \in \ec{x}$, where $z$ and $w$ are variables of integer type.
A possible application of the above rules, showing the state of $\Val$ after each application, is as follows:
\[\begin{array}{l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l@{\hspace{.5em}}l}
\hline
\text{Rule} & \multicolumn{4}{c}{\Val} \ 
\\
\hline
& \{ & \ec{x} \mapsto v_{\ec{x}}, & \ec{z} \mapsto v_{\ec{z}}, \\
&    & \ec{y} \mapsto v_{\ec{y}}, & \ec{w} \mapsto v_{\ec{w}} & \} \\
\rn{Ind} \ec{x}
& \{ & \ec{x} \mapsto \mu v_{\ec{x}}. \const{C}( v_{\ec{w}}, v_{\ec{y}} ), & \ec{z} \mapsto v_{\ec{z}}, \\
&    & \ec{y} \mapsto v_{\ec{y}}, & \ec{w} \mapsto v_{\ec{w}} & \} \\
\rn{Ind} \ec{y}
& \{ & \ec{x} \mapsto \mu v_{\ec{x}}. \const{C}( v_{\ec{w}}, \mu v_{\ec{y}}. \const{C}( v_{\ec{z}}, v_{\ec{x}} ) ), & \ec{z} \mapsto v_{\ec{z}}, \\
&    & \ec{y} \mapsto \mu v_{\ec{y}}. \const{C}( v_{\ec{z}}, v_{\ec{x}} ), & \ec{w} \mapsto v_{\ec{w}} & \} \\
\rn{Ind} \ec{x}
& \{ & \ec{x} \mapsto \mu v_{\ec{x}}. \const{C}( v_{\ec{w}}, \mu v_{\ec{y}}. \const{C}( v_{\ec{z}}, v_{\ec{x}} ) ), & \ec{z} \mapsto v_{\ec{z}}, \\
&    & \ec{y} \mapsto \mu v_{\ec{y}}. \const{C}( v_{\ec{z}}, \mu v_{\ec{x}}. \const{C}( v_{\ec{w}}, v_{\ec{y}} ) ), & \ec{w} \mapsto v_{\ec{w}} & \} \\
\rn{Base} \ec{z}
& \{ & \ec{x} \mapsto \mu v_{\ec{x}}. \const{C}( v_{\ec{w}}, \mu v_{\ec{y}}. \const{C}( \const{0}, v_{\ec{x}} ) ), & \ec{z} \mapsto \const{0}, \\
&    & \ec{y} \mapsto \mu v_{\ec{y}}. \const{C}( \const{0}, \mu v_{\ec{x}}. \const{C}( v_{\ec{w}}, v_{\ec{y}} ) ), & \ec{w} \mapsto v_{\ec{w}} & \} \\
\rn{Base} \ec{w}
& \{ & \ec{x} \mapsto \mu v_{\ec{x}}. \const{C}( \const{1}, \mu v_{\ec{y}}. \const{C}( \const{0}, v_{\ec{x}} ) ), & \ec{z} \mapsto \const{0}, \\
&    & \ec{y} \mapsto \mu v_{\ec{y}}. \const{C}( \const{0}, \mu v_{\ec{x}}. \const{C}( \const{1}, v_{\ec{y}} ) ), & \ec{w} \mapsto \const{1} & \} \\
\end{array}\]

It is clear that after applying these rules to completion, $\Val$ is a mapping from equivalence classes to ground values.
